---
title: "figures"
output: html_document
---

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(ggplot2)
library(showtext)
library(dplyr)
library(tidyboot)
library(lmerTest)
library(ggpubr)

showtext_auto()

font_add(family = "Lato", regular = here("fonts/Lato-Regular.ttf"))
theme_set(theme_classic(base_size = 15, base_family = "Lato"))
options(contrasts = c(unordered = "contr.sum", ordered = "contr.poly"))

action_colors <- c("#F0E442", "#E69F00", "#D6D6D6")
relationship_colors = c("Higher" = "#DC267F", "Lower" = "#785EF0")
```


# Study 1

```{r}
study_1_data <- read_csv(here("data/study-1_tidy_data.csv")) %>%
  mutate(
    relationship = factor(
      relationship,
      levels = c("No info", "Symmetric", "Asymmetric")
    ),
    next_interaction = factor(
      next_interaction,
      levels = c("Reciprocity", "Precedent", "None")
    )
  )

study_1_summary <- study_1_data %>%
  group_by(relationship, next_interaction) %>%
  tidyboot_mean(likert_rating, na.rm = T)

study_1_summary_scenario <- study_1_data %>%
  group_by(relationship, next_interaction, story) %>%
  tidyboot_mean(likert_rating, na.rm = T)
```


```{r fig.width=6, fig.height=3.5, dpi=125}
ggplot(study_1_data,
       aes(x = relationship, y = likert_rating, fill = next_interaction)) +
  geom_violin(bw = 0.43,
              position = position_dodge(width = 0.7),
              alpha = 0.8) +
  geom_pointrange(
    data = study_1_summary,
    aes(
      x = relationship,
      y = empirical_stat,
      ymin = ci_lower,
      ymax = ci_upper
    ),
    size = 0.3,
    linewidth = 1,
    position = position_dodge(width = 0.7)
  ) +
  scale_fill_manual(values = action_colors) +
  scale_color_manual(values = action_colors) +
  scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7)) +
  labs(title = "Study 1: Equal vs. hierarchical",
       x = "Relationship",
       y = "How likely?",
       fill = "Next interaction") +
  theme(legend.position = "bottom")

ggsave(here("figures/study-1_main.pdf"))
```

```{r fig.width=10, fig.height=10.5}
ggplot(study_1_data,
       aes(x = relationship, y = likert_rating, fill = next_interaction)) +
  geom_violin(bw = 0.43,
              position = position_dodge(width = 0.7),
              alpha = 0.8) +
  geom_pointrange(
    data = study_1_summary_scenario,
    aes(
      x = relationship,
      y = empirical_stat,
      ymin = ci_lower,
      ymax = ci_upper
    ),
    size = 0.3,
    linewidth = 1,
    position = position_dodge(width = 0.7)
  ) +
  scale_fill_manual(values = action_colors) +
  scale_color_manual(values = action_colors) +
  scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7)) +
  labs(title = "Study 1: Equal vs. hierarchical",
       x = "Relationship",
       y = "How likely?",
       fill = "Next interaction") +
  theme(legend.position = "bottom") +
  facet_wrap( ~ story, nrow = 6)

# ggsave(here("figures/study-1_scenarios.pdf"))
```

# Study 2

```{r}
study_2_data <- read_csv(here("data/study-2_tidy_data.csv")) %>% 
    mutate(
    relationship = factor(
      relationship,
      levels = c("Equal", "Lower", "Higher")
    ),
    next_interaction = factor(
      next_interaction,
      levels = c("Reciprocity", "Precedent", "None")
    )
  )

study_2_summary <- study_2_data %>%
  group_by(relationship, next_interaction) %>%
  tidyboot_mean(likert_rating, na.rm = T)

study_2_summary_scenario <- study_2_data %>%
  group_by(relationship, next_interaction, story) %>%
  tidyboot_mean(likert_rating, na.rm = T)
```


```{r fig.width=6, fig.height=3.5, dpi=125}
ggplot(study_2_data,
       aes(x = relationship, y = likert_rating, fill = next_interaction)) +
  geom_violin(bw = 0.43,
              position = position_dodge(width = 0.7),
              alpha = 0.8) +
  geom_pointrange(
    data = study_2_summary,
    aes(
      x = relationship,
      y = empirical_stat,
      ymin = ci_lower,
      ymax = ci_upper
    ),
    size = 0.3,
    linewidth = 1,
    position = position_dodge(width = 0.7)
  ) +
  scale_fill_manual(values = action_colors) +
  scale_color_manual(values = action_colors) +
  scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7)) +
  labs(title = "Study 2: Relative rank",
       x = "Rank of generous character",
       y = "How likely?",
       fill = "Next interaction") +
  theme(legend.position = "bottom")

# ggsave(here("figures/study-2_main.pdf"))
```

```{r fig.width=10, fig.height=10.5}
ggplot(study_2_data,
       aes(x = relationship, y = likert_rating, fill = next_interaction)) +
  geom_violin(bw = 0.43,
              position = position_dodge(width = 0.7),
              alpha = 0.8) +
  geom_pointrange(
    data = study_2_summary_scenario,
    aes(
      x = relationship,
      y = empirical_stat,
      ymin = ci_lower,
      ymax = ci_upper
    ),
    size = 0.3,
    linewidth = 1,
    position = position_dodge(width = 0.7)
  ) +
  scale_fill_manual(values = action_colors) +
  scale_color_manual(values = action_colors) +
  scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7)) +
  labs(title = "Study 2: Relative rank",
       x = "Rank of generous character",
       y = "How likely?",
       fill = "Next interaction") +
  theme(legend.position = "bottom") +
  facet_wrap( ~ story, nrow = 6)

# ggsave(here("figures/study-2_scenarios.pdf"))
```
## Validation study

TODO: Move from other R script

## Effect size benefit over effort

```{r}
study_2_benefit_effort <- read.csv(here("data/study-2_benefit_effort.csv")) %>% 
  rename(
    diff_x = diff,
    ci_lower_x = ci_lower, 
    ci_upper_x = ci_upper) %>% 
  mutate(relationship = factor(relationship, levels = c("Lower", "Higher")))
```

```{r}
study_2_benefit_effort_summary <- study_2_benefit_effort %>%
  group_by(story, relationship, type, diff_x, ci_lower_x, ci_upper_x) %>%
  tidyboot_mean(p_prec, na.rm = TRUE) %>% 
  rename(p_prec = empirical_stat)
```

```{r fig.width = 7, fig.height=4}
appender <- function(string, suffix = "-rank generous") paste0(string, suffix)

ggplot(study_2_benefit_effort_summary %>% filter(type == "benefit"), aes(x = diff_x, y = p_prec, color = relationship)) +
  geom_point(size = 3, alpha = 0.3) +
  geom_smooth(method = "lm", fill = "lightgray") +
  geom_errorbar(mapping = aes(x = diff_x, ymin = ci_lower, ymax=ci_upper), size= 1.5, width = 0.13, alpha = 0.3) +
    geom_errorbarh(mapping = aes(y = p_prec, xmin = ci_lower_x, xmax = ci_upper_x), size= 1.5, height = 0.015, alpha = 0.3) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  scale_color_manual(values = c("Higher" = "#DC267F", "Lower" = "#785EF0")) +
  labs(x = "Differential benefit",
       y = "P(Precedent)",
       title = "Study 2: Differential benefit vs. likelihood of a precedent") + 
  theme(legend.position = "none") +
  facet_grid(~ relationship, labeller = as_labeller(appender))

ggsave(here("figures/study-2_benefit.pdf"))
```
```{r fig.width = 7, fig.height=4}
ggplot(study_2_benefit_effort_summary %>% filter(type == "effort"), aes(x = diff_x, y = p_prec, color = relationship)) +
  geom_point(size = 3, alpha = 0.3) +
  geom_smooth(method = "lm", fill = "lightgray") +
  geom_errorbar(mapping = aes(x = diff_x, ymin = ci_lower, ymax=ci_upper), size= 1.5, width = 0.13, alpha = 0.3) +
    geom_errorbarh(mapping = aes(y = p_prec, xmin = ci_lower_x, xmax = ci_upper_x), size= 1.5, height = 0.015, alpha = 0.3) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
    scale_color_manual(values = c("Higher" = "#DC267F", "Lower" = "#785EF0")) +
  labs(x = "Differential effort",
       y = "P(Precedent)",
       title = "Study 2: Differential effort vs. likelihood of a precedent") + 
  theme(legend.position = "none") +
  facet_grid(~ relationship, labeller = as_labeller(appender))

ggsave(here("figures/study-2_effort.pdf"))
```

# Study 3

```{r}
study_3_data <- read_csv(here("data/study-3_tidy_data.csv")) %>% 
    mutate(
    relationship = factor(
      relationship,
      levels = c("Equal", "Lower", "Higher")
    ),
    next_interaction = factor(
      next_interaction,
      levels = c("Reciprocity", "Precedent", "None")
    )
  )

study_3_summary <- study_3_data %>%
  group_by(relationship, next_interaction) %>%
  tidyboot_mean(likert_rating, na.rm = T)

study_3_summary_scenario <- study_3_data %>%
  group_by(relationship, next_interaction, story) %>%
  tidyboot_mean(likert_rating, na.rm = T)
```


```{r fig.width=6, fig.height=3.5, dpi=125}
ggplot(study_3_data,
       aes(x = relationship, y = likert_rating, fill = next_interaction)) +
  geom_violin(bw = 0.43,
              position = position_dodge(width = 0.7),
              alpha = 0.8) +
  geom_pointrange(
    data = study_3_summary,
    aes(
      x = relationship,
      y = empirical_stat,
      ymin = ci_lower,
      ymax = ci_upper
    ),
    size = 0.3,
    linewidth = 1,
    position = position_dodge(width = 0.7)
  ) +
  scale_fill_manual(values = action_colors) +
  scale_color_manual(values = action_colors) +
  scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7)) +
  labs(title = "Study 3: Concrete relationships",
       x = "Rank of generous character",
       y = "How likely?",
       fill = "Next interaction") +
  theme(legend.position = "bottom")
# ggsave(here("figures/study-4_main.pdf"))
```

```{r fig.width=10, fig.height=10.5}
ggplot(study_3_data,
       aes(x = relationship, y = likert_rating, fill = next_interaction)) +
  geom_violin(bw = 0.43,
              position = position_dodge(width = 0.7),
              alpha = 0.8) +
  geom_pointrange(
    data = study_3_summary_scenario,
    aes(
      x = relationship,
      y = empirical_stat,
      ymin = ci_lower,
      ymax = ci_upper
    ),
    size = 0.3,
    linewidth = 1,
    position = position_dodge(width = 0.7)
  ) +
  scale_fill_manual(values = action_colors) +
  scale_color_manual(values = action_colors) +
  scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7)) +
  labs(title = "Study 3: Concrete relationships",
       x = "Rank of generous character",
       y = "How likely?",
       fill = "Next interaction") +
  theme(legend.position = "bottom") +
  facet_wrap( ~ story, nrow = 6)

ggsave(here("figures/study-3_scenarios.pdf"))
```


## Effect size benefit over effort

```{r}
study_3_benefit_effort <- read.csv(here("data/study-3_benefit_effort.csv")) %>% 
  rename(
    diff_x = diff,
    ci_lower_x = ci_lower, 
    ci_upper_x = ci_upper) %>% 
  mutate(relationship = factor(relationship, levels = c("Lower", "Higher")))
```

```{r}
study_3_benefit_effort_summary <- study_3_benefit_effort %>%
  group_by(story, relationship, type, diff_x, ci_lower_x, ci_upper_x) %>%
  tidyboot_mean(p_prec, na.rm = TRUE) %>% 
  rename(p_prec = empirical_stat)
```

```{r fig.width = 7, fig.height=4}
appender <- function(string, suffix = "-rank generous") paste0(string, suffix)

ggplot(study_3_benefit_effort_summary %>% filter(type == "benefit"), aes(x = diff_x, y = p_prec, color = relationship)) +
  geom_point(size = 3, alpha = 0.3) +
  geom_smooth(method = "lm", fill = "lightgray") +
  geom_errorbar(mapping = aes(x = diff_x, ymin = ci_lower, ymax=ci_upper), size= 1.5, width = 0.13, alpha = 0.3) +
    geom_errorbarh(mapping = aes(y = p_prec, xmin = ci_lower_x, xmax = ci_upper_x), size= 1.5, height = 0.015, alpha = 0.3) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  scale_color_manual(values = c("Higher" = "#DC267F", "Lower" = "#785EF0")) +
  labs(x = "Differential benefit",
       y = "P(Precedent)",
       title = "Study 3: Differential benefit vs. likelihood of a precedent") + 
  theme(legend.position = "none") +
  facet_grid(~ relationship, labeller = as_labeller(appender))

ggsave(here("figures/study-3_benefit.pdf"))
```
```{r fig.width = 7, fig.height=4}
ggplot(study_3_benefit_effort_summary %>% filter(type == "effort"), aes(x = diff_x, y = p_prec, color = relationship)) +
  geom_point(size = 3, alpha = 0.3) +
  geom_smooth(method = "lm", fill = "lightgray") +
  geom_errorbar(mapping = aes(x = diff_x, ymin = ci_lower, ymax=ci_upper), size= 1.5, width = 0.13, alpha = 0.3) +
    geom_errorbarh(mapping = aes(y = p_prec, xmin = ci_lower_x, xmax = ci_upper_x), size= 1.5, height = 0.015, alpha = 0.3) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
    scale_color_manual(values = c("Higher" = "#DC267F", "Lower" = "#785EF0")) +
  labs(x = "Differential effort",
       y = "P(Precedent)",
       title = "Study 3: Differential effort vs. likelihood of a precedent") + 
  theme(legend.position = "none") +
  facet_grid(~ relationship, labeller = as_labeller(appender))

ggsave(here("figures/study-3_effort.pdf"))
```

# Study 4

```{r}
study_4_data <- read_csv(here("data/study-4_tidy_data.csv")) %>%
  mutate(
    first_actual = factor(
      first_actual,
      levels = c("Equal", "Lower", "Higher")
    ),
    first_response = factor(
      first_response,
      levels = c("Equal", "Lower", "Higher")
    )
  )

# Implicit coordination
study_4_first_time_summary <-
  study_4_data %>%
  mutate(first_response = recode(first_response, "Higher" = 0, "Lower" = 1)) %>%
  group_by(story, diff_effort, ci_lower_effort, ci_upper_effort, diff_benefit, ci_lower_benefit, ci_upper_benefit) %>%
  tidyboot_mean(first_response, na.rm = TRUE)


d.temp <- study_4_first_time_summary %>%
  arrange(desc(mean))

levs <- unique(d.temp$story)

study_4_first_time_summary$story <-
  factor(study_4_first_time_summary$story, levels = levs)


# First vs. second time
study_4_second_summary <-
  study_4_data %>% filter(symmetric == "Asymmetric") %>%
  mutate(second_response = recode(second_response, "Higher" = 0, "Lower" = 1)) %>%
  group_by(story, first_actual) %>%
  tidyboot_mean(second_response, na.rm = TRUE)

study_4_first_second_summary <-
  left_join(
    study_4_first_time,
    study_4_second_time,
    suffix = c("_first", "_second"),
    by = (c("story"))
  )
```
## Do we replicate the previous results? 

```{r}
study_4_summary <- study_4_data %>% 
    mutate(strategy = recode(
    strategy,
    "Precedent" = 1,
    "Reciprocity" = 0
  )) %>%
  group_by(first_response, first_actual) %>% 
  tidyboot_mean(strategy, na.rm = TRUE) %>% 
  mutate(Expectations = case_when(
  first_actual != first_response ~ "Inconsistent",
  first_actual == first_response ~ "Consistent"), 
  Expectations = factor(Expectations, levels = c("Consistent", "Inconsistent")))
```

      

```{r fig.width=6, fig.height=3.2}
ggplot(study_4_summary, aes(x = first_actual, y = empirical_stat, color = first_actual, shape = Expectations)) +
  geom_pointrange(
    aes(ymin = ci_lower, ymax = ci_upper),
    size = 0.7,
    linewidth = 1.5,
    position = position_dodge(width = 0.2),
  ) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray") +
  scale_color_manual(values = relationship_colors, guide = "none") +
  scale_shape_discrete(name = "First time prediction") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Rank of generous actor",
       y = "P(Precedent)") +
  theme(legend.position = "right") 
```

## Implicit cooordination

First time plots

```{r fig.width=8, fig.height=3.5, dpi=125}
ggplot(study_4_first_time_summary, aes(x = story, y = empirical_stat)) +
  geom_pointrange(
    data = study_4_first_time_summary,
    aes(
      x = story,
      y = empirical_stat,
      ymin = ci_lower,
      ymax = ci_upper
    ),
    size = 0.7,
    linewidth = 1.5
  ) +
  geom_hline(
    yintercept = 0.5,
    linetype = "dashed",
    color = "gray",
    alpha = 0.5
  ) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Scenario", y = "P(lower-rank generous)", title = 'Study 3: Implicit coordination') +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

ggsave(here("figures/study-4_implicit.pdf"))
```

## Main plot: First time vs. second time

```{r fig.width=4.5, fig.height=4.5}
ggplot(
    study_4_first_second_summary,
    aes(x = empirical_stat_first, y = empirical_stat_second, color = first_actual)
  ) +
  geom_smooth(method = "lm",
              fill = "lightgray",
              linewidth = 1.3) +
  geom_point(size = 3.6,
             alpha = 0.3,
             stroke = 0) +
  geom_errorbar(
    mapping = aes(x = empirical_stat_first, ymin = ci_lower_second, ymax = ci_upper_second),
    size = 1.5,
    width = 0.03,
    alpha = 0.3
  ) +
  geom_errorbarh(
    mapping = aes(y = empirical_stat_second, xmin = ci_lower_first, xmax = ci_upper_first),
    size = 1.5,
    height = 0.03,
    alpha = 0.3
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  geom_vline(xintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(0, 1)) +
  scale_color_manual(values = c("Higher" = "#DC267F", "Lower" = "#785EF0")) +
  labs(x = "First time P(lower-rank generous)", y = "Second time P(lower-rank generous)", color = "Observed rank of generous actor") + 
  theme(legend.position = "bottom") 



ggsave(here("figures/study-4_first_second.pdf"))
```

## Study 4 vs. Study 2

```{r}

study_2_first_second <- study_2_data %>%
  filter(relationship != "Equal", next_interaction != 'None') %>%
  group_by(subject_id, story, relationship) %>%
  mutate(
    total_rating = sum(likert_rating),
    normalized_likert_rating = likert_rating / total_rating
  ) %>%
  select(-total_rating) %>%
  ungroup() %>%
  rename(first_actual  = relationship) %>%
  mutate(
    second_response = case_when(
      next_interaction == 'Precedent' ~ first_actual,
      next_interaction == 'Reciprocity' ~ ifelse(first_actual == "Higher", "Lower", "Higher"),
      next_interaction == 'None' ~ 'None'
    )
  )


study_2_first_second_summary <- study_2_first_second %>%
  filter(second_response == 'Lower') %>%
  group_by(story, first_actual) %>%
  tidyboot_mean(normalized_likert_rating, na.rm = T)


studies_2_4_summary <-
  left_join(
    study_2_first_second_summary,
    study_4_first_time_summary,
    suffix = c("_2", "_4"),
    by = (c("story"))
  )

studies_2_4_all <- 
    left_join(
    study_2_first_second,
    study_4_first_time,
    suffix = c("_2", "_4"),
    by = (c("story", "diff_benefit", "ci_lower_benefit", "ci_upper_benefit", "diff_effort", "ci_lower_effort", "ci_upper_effort"))
  ) %>% 
  rename(expected.first.4 = empirical_stat, expected.next.2 = normalized_likert_rating) %>%
  filter(second_response == "Lower")


```


```{r fig.width=4.5, fig.height=4.5}
ggplot(studies_2_4_summary,
       aes(x = empirical_stat_4, y = empirical_stat_2, color = first_actual)) +
  geom_point(size = 3.3,
             alpha = 0.3,
             stroke = 0) +
  geom_smooth(method = "lm",
              fill = "lightgray",
              linewidth = 1.3) +
  geom_errorbar(
    mapping = aes(x = empirical_stat_4, ymin = ci_lower_2, ymax = ci_upper_2),
    size = 1.5,
    width = 0.03,
    alpha = 0.3
  ) +
  geom_errorbarh(
    mapping = aes(y = empirical_stat_2, xmin = ci_lower_4, xmax = ci_upper_4),
    size = 1.5,
    height = 0.03,
    alpha = 0.3
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  geom_vline(xintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0.2, 0.8)) +
  scale_x_continuous(limits = c(0, 1)) +
  scale_color_manual(values = c("Higher" = "#DC267F", "Lower" = "#785EF0")) +
  labs(x = "Study 4 First time P(lower-rank generous)", y = "Study 2 P(lower-rank generous)", color = "Observed rank of generous actor") +
  theme(legend.position = "bottom")


ggsave(here("figures/study-2-4_first_second.pdf"))
```

## Study 4 vs. study 3 (concrete relationships) 

```{r}
study_3_first_second <- study_3_data %>%
  filter(relationship != "Equal", next_interaction != 'None') %>%
  group_by(subject_id, story, relationship) %>%
  mutate(
    total_rating = sum(likert_rating),
    normalized_likert_rating = likert_rating / total_rating
  ) %>%
  select(-total_rating) %>%
  ungroup() %>%
  rename(first_actual = relationship) %>%
  mutate(
    second_response = case_when(
      next_interaction == 'Precedent' ~ first_actual,
      next_interaction == 'Reciprocity' ~ ifelse(first_actual == "Higher", "Lower", "Higher"),
      next_interaction == 'None' ~ 'None'
    )
  )

study_3_first_second_summary <- study_3_first_second %>%
  filter(second_response == 'Lower') %>%
  group_by(story, first_actual, prestige_score, dominance_score) %>%
  tidyboot_mean(normalized_likert_rating, na.rm = T)


studies_3_4_summary <-
  left_join(
    study_3_first_second_summary,
    study_4_first_time_summary,
    suffix = c("_3", "_4"),
    by = (c("story"))
  )

studies_3_4_all <- 
    left_join(
    study_3_first_second,
    study_4_first_time,
    suffix = c("_3", "_4"),
    by = (c("story", "diff_benefit", "ci_lower_benefit", "ci_upper_benefit", "diff_effort", "ci_lower_effort", "ci_upper_effort"))
  ) %>% 
  rename(expected.first.3 = empirical_stat, expected.next.4 = normalized_likert_rating) %>%
  filter(second_response == "Lower")
```


```{r fig.width=4.5, fig.height=4.5}
ggplot(studies_3_4_summary,
       aes(x = empirical_stat_4, y = empirical_stat_3, color = first_actual)) +
  geom_point(size = 3.3,
             alpha = 0.3,
             stroke = 0) +
  geom_smooth(method = "lm",
              fill = "lightgray",
              linewidth = 1.3) +
  geom_errorbar(
    mapping = aes(x = empirical_stat_4, ymin = ci_lower_3, ymax = ci_upper_3),
    size = 1.5,
    width = 0.03,
    alpha = 0.3
  ) +
  geom_errorbarh(
    mapping = aes(y = empirical_stat_3, xmin = ci_lower_4, xmax = ci_upper_4),
    size = 1.5,
    height = 0.03,
    alpha = 0.3
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  geom_vline(xintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0.2, 0.8)) +
  scale_x_continuous(limits = c(0, 1)) +
  scale_color_manual(values = c("Higher" = "#DC267F", "Lower" = "#785EF0")) +
  labs(x = "Study 4 First time P(lower-rank generous)", y = "Study 3 P(lower-rank generous)", color = "Observed rank of generous actor") +
  theme(legend.position = "bottom")


ggsave(here("figures/study-3-4_first_second.pdf"))
```

```{r fig.width=7.5, fig.height=4.5}
ggplot(studies_3_4_summary,
       aes(x = prestige_score, y = empirical_stat_3, color = first_actual)) +
  geom_point(size = 3.3,
             alpha = 0.3,
             stroke = 0) +
  geom_smooth(method = "lm",
              fill = "lightgray",
              linewidth = 1.3) +
  geom_errorbar(
    mapping = aes(x = prestige_score, ymin = ci_lower_3, ymax = ci_upper_3),
    size = 1.5,
    width = 0.03,
    alpha = 0.3
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0.2, 0.8)) +
  scale_color_manual(values = c("Higher" = "#DC267F", "Lower" = "#785EF0")) +
  labs(x = "Prestige score", y = "Study 3 P(lower-rank generous)", color = "Observed rank of generous actor") +
  theme(legend.position = "right")


ggsave(here("figures/study-3_prestige.pdf"))
```

```{r fig.width=7.5, fig.height=4.5}
ggplot(studies_3_4_summary,
       aes(x = dominance_score, y = empirical_stat_3, color = first_actual)) +
  geom_point(size = 3.3,
             alpha = 0.3,
             stroke = 0) +
  geom_smooth(method = "lm",
              fill = "lightgray",
              linewidth = 1.3) +
  geom_errorbar(
    mapping = aes(x = dominance_score, ymin = ci_lower_3, ymax = ci_upper_3),
    size = 1.5,
    width = 0.03,
    alpha = 0.3
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0.2, 0.8)) +
  scale_color_manual(values = c("Higher" = "#DC267F", "Lower" = "#785EF0")) +
  labs(x = "Dominance score", y = "Study 3 P(lower-rank generous)", color = "Observed rank of generous actor") +
  theme(legend.position = "right")


ggsave(here("figures/study-3_dominance.pdf"))
```

### Regression

```{r}
studies_4_3_all <- studies_4_3_all %>% 
  rename(expected.next.3 = expected.next.4, 
         expected.first.4 = expected.first.3)
```


Same regression as before

```{r}
mod1 <- lmer(
  data = studies_4_3_all,
  expected.next.3 ~ expected.first.4 * observed_higher + (1 |
                                                              subject_id) + (1 | story)
)

summary(mod1)
```
Regression with dominance and prestige score

```{r}
mod2 <- lmer(
  data = studies_4_3_all,
  expected.next.3 ~ (dominance_score + prestige_score + expected.first.4) * observed_higher  + (1 |
                                                              subject_id) + (1 | story)
)

summary(mod2)
```
```{r}
anova(mod1, mod2)
```
plot correlation of dominance score vs. prestige score

```{r}
ggplot(studies_4_3_all, aes(x = prestige_score, y = dominance_score)) +
  geom_point() +
  labs(x = "Prestige score", y = "Dominance score") 
```

Plot 'first time' with dominance score, see how much they are related

```{r fig.width=4.5, fig.height=4.5}
# TODO: add CIs
ggplot(studies_3_4,
       aes(x = empirical_stat_4, y = dominance_score)) +
  geom_point(size = 3.3,
             alpha = .8,
             stroke = 0) +
  geom_smooth(method = "lm",
              fill = "lightgray",
              linewidth = 1.3) +
  labs(x = "Study 4 First time P(lower-rank generous)", y = "Dominance score") 

ggplot(studies_3_4,
       aes(x = empirical_stat_4, y = prestige_score)) +
  geom_point(size = 3.3,
             alpha = .8,
             stroke = 0) +
  geom_smooth(method = "lm",
              fill = "lightgray",
              linewidth = 1.3) +
  labs(x = "Study 4 First time P(lower-rank generous)", y = "Prestige score") 


```



# Study 5

```{r}
study_5_data <- read_csv(here("data/study-5_tidy_data.csv")) %>%
  mutate(
    participant_rank = case_when(
      partner_status == "equal" ~ "Equal",
      partner_status == "lower" ~ "Higher",
      partner_status == "higher" ~ "Lower"
    ),
    participant_rank = factor(participant_rank, levels = c("Equal", "Lower", "Higher")),
    participant_first_choice = factor(
      participant_first_choice,
      levels = c("receive", "give"),
      labels = c("Receive", "Give")
    ),
    partner_first_choice = factor(
      partner_first_choice,
      levels = c("receive", "give"),
      labels = c("Receive", "Give")
    )
  )

# Analyze first time choices by partner status
study_5_first_time_summary <- study_5_data %>%
  group_by(scenario_id, participant_rank) %>%
  tidyboot_mean(participant_first_choice_generous, na.rm = T)

study_5_first_time_all <- study_5_data %>%
  group_by(participant_rank) %>%
  tidyboot_mean(participant_first_choice_generous, na.rm = T)
```

```{r fig.width=6, fig.height=4}
ggplot(
  study_5_first_time_summary,
  aes(x = participant_rank, y = empirical_stat)
) +
  geom_pointrange(
    aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper),
    size = 0.7,
    linewidth = 1.5
  ) +
  labs(
    title = "Study 5: First time",
    x = "Self rank",
    y = "P(first time generous)"
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_manual(values = relationship.colors) +
  facet_wrap( ~ scenario_id) +
  theme(legend.position = "none")

ggsave(here("figures/study-5_first_time_scenarios.pdf"))
```

```{r fig.width=2.8, fig.height=3.2}
ggplot(
  study_5_first_time_all,
  aes(x = participant_rank, y = empirical_stat)
) +
  geom_pointrange(
    aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper),
    size = 0.7,
    linewidth = 1.5
  ) +
  labs(
    title = "Study 5: First time",
    x = "Self rank",
    y = "P(Self generous first time)",
    color = "Self rank"
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_manual(values = relationship.colors) +
  theme(legend.position = "none")

ggsave(here("figures/study-5_first_time_all.pdf"))
```


```{r}
study_5_strategy_summary <- study_5_data %>%
  group_by(participant_rank, successful_first_time) %>%
  tidyboot_mean(strategy_alternating, na.rm = T)
```

```{r fig.width=5.5, fig.height=3.2}
ggplot(
  study_5_strategy_summary,
  aes(
    x = participant_rank,
    y = empirical_stat,
    color = factor(successful_first_time)
  )
) +
  geom_pointrange(
    aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper),
    size = 0.7,
    linewidth = 1.5,
    position = position_dodge(width = 0.2)
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  labs(
    title = "Study 5: Strategies",
    x = "Self rank",
    y = "P(Reciprocity)",
    color = "First time coordination"
  ) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_brewer(palette = "Set1", labels = c("Unsuccessful", "Successful")) 

ggsave(here("figures/study-5_strategies.pdf"))
```

```{r}
study_5_summary <- study_5_data %>%
  group_by(
    participant_rank,
    participant_first_choice,
    partner_first_choice,
    successful_first_time
  ) %>%
  tidyboot_mean(participant_second_choice_generous, na.rm = TRUE)
```

```{r fig.width = 4.6, fig.height = 5}
ggplot(
  study_5_summary,
  aes(x = participant_rank, y = empirical_stat, color = successful_first_time)
) +
  # Add shaded rectangles based on partner_first_choice
  geom_rect(
    data = data.frame(
      partner_first_choice = c("Receive", "Receive", "Give", "Give"),
      ymin = c(0.5, 0, 0, 0.5),
      ymax = c(1, 0.5, 0.5, 1),
      rect_color = c("#E69F00", "#F0E442", "#E69F00", "#F0E442")
    ),
    aes(
      xmin = -Inf,
      xmax = Inf,
      ymin = ymin,
      ymax = ymax,
      fill = rect_color
    ),
    alpha = 0.15,
    inherit.aes = FALSE
  ) +
  geom_pointrange(
    aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper),
    size = 0.7,
    linewidth = 1.5,
    position = position_dodge(width = 0.3)
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = c("#E69F00" =  "#E69F00", "#F0E442" = "#F0E442")) +
  labs(
    title = "Study 5: Second time",
    x = "Self rank",
    y = "P(Second time generous)",
    color = "Self rank"
  ) +
  scale_color_brewer(
    palette = "Set1", 
    labels = c("Unsuccessful", "Successful")
  ) + 
  facet_grid(partner_first_choice ~ participant_first_choice) +
  theme(legend.position = "none", strip.background = element_blank())

ggsave(here("figures/study-5_second_time.pdf"))
```
Plot coordination and miscoordination separately

```{r fig.width=5.1, fig.height=3.2}
ggplot(study_5_summary %>% filter(successful_first_time == TRUE), aes(x = participant_rank, y = empirical_stat, color = partner_first_choice, shape = partner_first_choice)) +
  geom_pointrange(
    aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper),
    size = 0.7,
    linewidth = 1.5,
    position = position_dodge(width = 0.3)
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    title = "Study 5: Second time (successful first time)",
    x = "Self rank",
    y = "P(Self generous second time)",
    color = "First time generous",
    shape = "First time generous"
  ) +
  scale_color_manual(
    values = c("Receive" = "#E69F00", "Give" = "#F0E442"), 
    labels = c("Receive" = "Self (Precedent)", "Give" = "Partner (Reciprocity)")
  ) + 
  scale_shape_manual(
    values = c("Receive" = 16, "Give" = 17),
    labels = c("Receive" = "Self (Precedent)", "Give" = "Partner (Reciprocity)")
  )

ggsave(here("figures/study-5_second_time_successful.pdf"))
```
```{r fig.width=4.5, fig.height=3.2}
ggplot(study_5_summary %>% filter(successful_first_time == FALSE), aes(x = participant_rank, y = empirical_stat, color = partner_first_choice)) +
  geom_pointrange(
    aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper),
    size = 0.7,
    linewidth = 1.5,
    position = position_dodge(width = 0.3)
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    title = "Study 5: Second time (unsuccessful first time)",
    x = "Self rank",
    y = "P(Self generous second time)",
    color = "First time"
  ) +
  scale_color_grey(
    start = 0.4, end = 0.7,
     labels = c("Give" = "Both give", "Receive" = "Both receive")
  ) 

ggsave(here("figures/study-5_second_time_unsuccessful.pdf"))
```


```{r}
ggplot(
  study_5_summary,
  aes(x = partner_first_choice, y = empirical_stat, color = participant_rank)
) +
  geom_pointrange(
    aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper),
    size = 0.7,
    linewidth = 1.5,
    position = position_dodge(width = 0.3)
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    title = "Study 5: Second time choices",
    x = "Observed partner first choice",
    y = "P(Second time generous)",
    color = "Self rank"
  ) +
  scale_color_manual(values = relationship.colors) +
  facet_wrap(~ participant_first_choice,
             labeller = labeller(
               participant_first_choice = function(x)
                 paste("First choice:", x)
             )) +
  theme(legend.position = "bottom")
```

# Study 6

```{r}
study_6_data <- read_csv(here("data/study-6_tidy_data.csv")) %>%
  mutate(
    participant_rank = case_when(
      partner_status == "equal" ~ "Equal",
      partner_status == "lower" ~ "Higher",
      partner_status == "higher" ~ "Lower"
    ),
    participant_rank = factor(participant_rank, levels = c("Equal", "Lower", "Higher")),
    participant_first_choice = factor(
      participant_first_choice,
      levels = c("receive", "give"),
      labels = c("Receive", "Give")
    ),
    partner_first_choice = factor(
      partner_first_choice,
      levels = c("receive", "give"),
      labels = c("Receive", "Give")
    )
  )

# Analyze first time choices by partner status
study_6_first_time_summary <- study_6_data %>%
  group_by(scenario_id, participant_rank) %>%
  tidyboot_mean(participant_first_choice_generous, na.rm = T)

study_6_first_time_all <- study_6_data %>%
  group_by(participant_rank) %>%
  tidyboot_mean(participant_first_choice_generous, na.rm = T)
```

```{r fig.width=6, fig.height=4}
ggplot(
  study_6_first_time_summary,
  aes(x = participant_rank, y = empirical_stat)
) +
  geom_pointrange(
    aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper),
    size = 0.7,
    linewidth = 1.5
  ) +
  labs(
    title = "Study 6: First time",
    x = "Self rank",
    y = "P(First time generous)",
    color = "Self rank"
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  facet_wrap( ~ scenario_id) +
  theme(legend.position = "none")

ggsave(here("figures/study-6_first_time_scenarios.pdf"))
```

```{r fig.width=2.8, fig.height=3.2}
ggplot(
  study_6_first_time_all,
  aes(x = participant_rank, y = empirical_stat)
) +
  geom_pointrange(
    aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper),
    size = 0.7,
    linewidth = 1.5
  ) +
  labs(
    title = "Study 6: First time",
    x = "Self rank",
    y = "P(Self generous first time)",
    color = "Self rank"
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  theme(legend.position = "none")

ggsave(here("figures/study-6_first_time_all.pdf"))
```


```{r}
study_6_strategy_summary <- study_6_data %>%
  group_by(participant_rank, successful_first_time) %>%
  tidyboot_mean(strategy_alternating, na.rm = T)
```

```{r fig.width=5.5, fig.height=3.2}
ggplot(
  study_6_strategy_summary,
  aes(
    x = participant_rank,
    y = empirical_stat,
    color = factor(successful_first_time)
  )
) +
  geom_pointrange(
    aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper),
    size = 0.7,
    linewidth = 1.5,
    position = position_dodge(width = 0.2)
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  labs(
    title = "Study 6: Strategies",
    x = "Self rank",
    y = "P(Reciprocity)",
    color = "First time coordination"
  ) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_brewer(palette = "Set1", labels = c("Unsuccessful", "Successful")) 

ggsave(here("figures/study-6_strategies.pdf"))
```

```{r}
study_6_summary <- study_6_data %>%
  group_by(
    participant_rank,
    participant_first_choice,
    partner_first_choice,
    successful_first_time
  ) %>%
  tidyboot_mean(participant_second_choice_generous, na.rm = TRUE)
```

```{r fig.width = 4.6, fig.height = 5}
ggplot(
  study_6_summary,
  aes(x = participant_rank, y = empirical_stat, color = successful_first_time)
) +
  # Add shaded rectangles based on partner_first_choice
  geom_rect(
    data = data.frame(
      partner_first_choice = c("Receive", "Receive", "Give", "Give"),
      ymin = c(0.5, 0, 0, 0.5),
      ymax = c(1, 0.5, 0.5, 1),
      rect_color = c("#E69F00", "#F0E442", "#E69F00", "#F0E442")
    ),
    aes(
      xmin = -Inf,
      xmax = Inf,
      ymin = ymin,
      ymax = ymax,
      fill = rect_color
    ),
    alpha = 0.15,
    inherit.aes = FALSE
  ) +
  geom_pointrange(
    aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper),
    size = 0.7,
    linewidth = 1.5,
    position = position_dodge(width = 0.3)
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = c("#E69F00" =  "#E69F00", "#F0E442" = "#F0E442")) +
  labs(
    title = "Study 6: Second time",
    x = "Self rank",
    y = "P(Second time generous)",
    color = "Self rank"
  ) +
  scale_color_brewer(
    palette = "Set1", 
    labels = c("Unsuccessful", "Successful")
  ) +
  facet_grid(partner_first_choice ~ participant_first_choice) +
  theme(legend.position = "none", strip.background = element_blank())

ggsave(here("figures/study-6_second_time.pdf"))
```

Plot coordination and miscoordination separately

```{r fig.width=5.1, fig.height=3.2}
ggplot(study_6_summary %>% filter(successful_first_time == TRUE), aes(x = participant_rank, y = empirical_stat, color = partner_first_choice)) +
  geom_pointrange(
    aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper),
    size = 0.7,
    linewidth = 1.5,
    position = position_dodge(width = 0.3)
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    title = "Study 6: Second time (successful first time)",
    x = "Self rank",
    y = "P(Self generous second time)",
    color = "First time generous"
  ) +
  scale_color_manual(
    values = c("Receive" = "#E69F00", "Give" = "#F0E442"), 
    labels = c("Receive" = "Self (Precedent)", "Give" = "Partner (Reciprocity)")
  )

ggsave(here("figures/study-6_second_time_successful.pdf"))
```

```{r fig.width=4.5, fig.height=3.2}
ggplot(study_6_summary %>% filter(successful_first_time == FALSE), aes(x = participant_rank, y = empirical_stat, color = partner_first_choice)) +
  geom_pointrange(
    aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper),
    size = 0.7,
    linewidth = 1.5,
    position = position_dodge(width = 0.3)
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    title = "Study 6: Second time (unsuccessful first time)",
    x = "Self rank",
    y = "P(Self generous second time)",
    color = "First time"
  ) +
  scale_color_grey(
    start = 0.4, end = 0.7,
     labels = c("Give" = "Both give", "Receive" = "Both receive")
  ) 

ggsave(here("figures/study-6_second_time_unsuccessful.pdf"))
```

```{r}
ggplot(
  study_6_summary,
  aes(x = partner_first_choice, y = empirical_stat, color = participant_rank)
) +
  geom_pointrange(
    aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper),
    size = 0.7,
    linewidth = 1.5,
    position = position_dodge(width = 0.3)
  ) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             color = "gray") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    title = "Study 6: Second time",
    x = "Observed partner first choice",
    y = "P(Second time generous)",
    color = "Self rank"
  ) +
  scale_color_manual(values = relationship.colors) +
  facet_wrap(~ participant_first_choice,
             labeller = labeller(
               participant_first_choice = function(x)
                 paste("First choice:", x)
             )) +
  theme(legend.position = "bottom")
```

# Cost/benefit validation
